A Survey of Physical Design Techniques of Information Systems

### 2010-07-05 cs848-db/Paper.tex
modified makefile to work with texmaker, just change the "Quick Build" user defined command to "make" and voila ;)
added some text to the data warehouses section, references

Note: when you compile the code, u will find in ur local repo a file named Paper.log (unversioned), use the context menu to ignore it :)

\subsection{Data Warehouses}
\label{SEC-WAREHOUSES}
\subsubsection*{Background}
Data warehouses, in contrast, are targeted for decision support. Historical, summarized and consolidated data is more important than detailed, individual records.

@@

bases for decision support systems \cite{248616}
OLAP vs OLTP
ROLAP
          o Data is stored in relational databases.
          o Support extensions to SQL.
          o Efficiently implement the multidimensional data model and operations.
          o Ex: Microstrategy.
MOLAP
          o Directly store multidimensional data in special data structures.
          o Implement the OLAP operations over these special data structures.
          o Star Schema.
          o Snowflake Schema.
          o Fact Constellations.
          o Ex: Essbase (Arbor), statistical DBMSes.
star schema
fact table

@@

\subsubsection*{Challenges}
- joins in data warehouses are usualy expensive since the fact table (the largest one) participates in every join \cite{628286}
-

\subsubsection*{Indexes}
\paragraph{Tree-based Indexes}
B-tree (B+ trees but commonly referred to as B-trees) \cite{253268}, R-tree \cite{602266}, K-D-B-tree \cite{582321}, BV-tree \cite{223796}

@@

\subsubsection*{Partitioning}
\paragraph{Vertical Partitioning}

@@

\subsubsection*{Data Cubes}

>>>>

\subsection{Data Warehouses}
\label{SEC-WAREHOUSES}
\subsubsection*{Background}

OLAP vs OLTP

OLTP \cite{sen2005comparison}

bases for decision support systems \cite{248616}
Data warehouses, in contrast, are targeted for decision support. Historical, summarized and consolidated data is more important than detailed, individual records.

@@

star schema
fact table

@@

- joins in data warehouses are usualy expensive since the fact table (the largest one) participates in every join \cite{628286}

\subsubsection*{Indexes}
\paragraph{Tree-based Indexes}
B-tree (B+ trees but commonly referred to as B-trees) \cite{253268}, R-tree \cite{602266, Cheung20011}, K-D-B-tree \cite{582321}, BV-tree \cite{223796}, UB-tree \cite{bayer1997universal}

@@

\subsubsection*{Partitioning}
\cite{thusoo2010data}
\paragraph{Vertical Partitioning}

@@

\subsubsection*{Data Cubes}
\cite{Cheung20011}
ROLAP
          o Data is stored in relational databases.
          o Support extensions to SQL.
          o Efficiently implement the multidimensional data model and operations.
          o Ex: Microstrategy.
MOLAP
          o Directly store multidimensional data in special data structures (arrays).
          o Implement the OLAP operations over these special data structures.
          o Star Schema.
          o Snowflake Schema. \cite{kimball2009data}
          o Fact Constellations.
          o Ex: Essbase (Arbor), statistical DBMSes.

hierachical data clusters \cite{karayannidis2008hierarchical}
